---
title: "Practical Machine Learning Course Project"
author: "Gabybs"
date: "July 26, 2015"
output: html_document
---

# Practical Machine Learning Course Project

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Reproducibility 
Loading packages and setting seed:
```{r, echo=TRUE}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)

set.seed(1234)
```

Loading the data:
```{r, echo=TRUE, results='hide'}
training <- read.csv("C:/Users/Gaby/Documents/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))

testing <- read.csv('C:/Users/Gaby/Documents/pml-testing.csv', na.strings=c("NA","#DIV/0!", ""))

# Deleting the columns with the missing values
training<-training[,colSums(is.na(training)) == 0]

testing <-testing[,colSums(is.na(testing)) == 0]

# Some variables are irrelevant to our current project, so we can delete them.
training <-training[,-c(1:7)]

testing <-testing[,-c(1:7)]

# New data set overlook:
dim(training)
dim(testing)
head(training)
head(testing)
```

## Cross-validation
In order to perform cross-validation, the training data set is partionned into 2 sets: subTraining (75%) and subTest (25%).

```{r, echo=TRUE, results='hide'}
subsamples <- createDataPartition(y=training$classe, p=0.75, list=FALSE)

subTraining <- training[subsamples, ] 
subTesting <- training[-subsamples, ]

head(subTraining)
head(subTesting)
```

The variable "classe" contains 5 levels: A, B, C, D and E. A plot of the outcome variable will allow us to see the frequency of each level.

```{r, echo=TRUE}
plot(subTraining$classe, col="red", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")
```

### First prediction model
```{r, echo=TRUE}
model1 <- rpart(classe ~ ., data=subTraining, method="class")

prediction1 <- predict(model1, subTesting, type = "class")

rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```

Test results:
```{r, echo=TRUE}
confusionMatrix(prediction1, subTesting$classe)
```

### Second prediction model
```{r, echo=TRUE}
model2 <- randomForest(classe ~. , data=subTraining, method="class")

prediction2 <- predict(model2, subTesting, type = "class")

# Test results:
confusionMatrix(prediction2, subTesting$classe)
```

According to the results, the Random forrest algorithm had a better performance. The accuracy from Random forrest is 95%, in comparison with the Tree Model that has a 73% accuracy. Wich is why we choose the random forrest algorithm.

Submission
```{r, echo=TRUE}
predictfinal <- predict(model2, testing, type="class")
predictfinal
```

```{r, echo=TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictfinal)
```
